---
title: "p8105_hw3_hj2660"
author: "Hyun Jin Jung"
date: "2023-10-14"
output: github_document
---
```{r setup, include=FALSE}
library(tidyverse)
library(tidyr)
library(knitr)

library(p8105.datasets)

knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  fig.width = 8,
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1

First, load the data from the `p8105.datasets`
```{r}
data(instacart)
```


The dataset includes `r nrow(instacart)` observations and `r ncol(instacart)` variables. Some of the key variables in the dataset include `order_id`, `product_id`, `user_id`, `aisle_id`, `department_id`, `product_name`, `order_dow`, and `order_hour_of_day`. One illustrative example is `r instacart[1,]$product_name` (Product ID: `r instacart[1,]$product_id`) found in the `r instacart[1,]$aisle` aisle in `r instacart[1,]$department` department. This product was purchased by the User ID: `r instacart[1,]$user_id` on the `r instacart[1,]$order_dow`th day of the week at `r instacart[1,]$order_hour_of_day`AM. It has been `r instacart[1,]$days_since_prior_order` days since prior order.


- How many aisles are there, and which aisles are the most items ordered from?

```{r}
instacart_aisle = instacart |>
  group_by(aisle) |>
  summarize(n_orders = n()) |>
  arrange(desc(n_orders))
```

There are `r nrow(instacart_aisle)` aisles. The aisle with the most items ordered is "`r instacart_aisle$aisle[1]`" with `r instacart_aisle$n_orders[1]` items.

- Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.

```{r}
instacart |>
  count(aisle) |>
  filter(n > 10000) |>
  mutate(aisle = fct_reorder(aisle, n)) |>
  ggplot(aes(x = aisle, y = n)) +
  geom_point() +
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

- Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.
```{r}
instacart |>
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) |>
  group_by(aisle) |>
  count(product_name) |>
  mutate(rank = min_rank(desc(n))) |>
  filter(rank < 4) |>
  arrange(desc(n)) |>
  knitr::kable()
```

- Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).
```{r}
instacart |>
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) |>
  group_by(product_name, order_dow) |>
  summarize(mean_hour = mean(order_hour_of_day)) |>
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour) |>
  knitr::kable(digits = 2)
```


## Problem 2

First, import dataset.
```{r}
data("brfss_smart2010")
```

Do some data cleaning, with a specific focus on the `Overall Health` topic. Include only responses that are ordered from `Poor` to `Excellent`.
```{r}
brfss_df = brfss_smart2010 |>
  janitor::clean_names() |>
  filter(topic == "Overall Health", response %in% c("Poor", "Fair", "Good", "Very good", "Excellent")) |>
  mutate(response = factor(response, levels = c("Poor", "Fair", "Good", "Very good", "Excellent")))
```


```{r}
states_2002 =
  brfss_df |>
  filter(year == "2002") |> 
  group_by(locationabbr) |>
  summarize(locations = n_distinct(locationdesc)) |>
  filter(locations >= 7)
kable(states_2002, col.names = c("States", "Number of Locations"))
```
In 2002, there are `r nrow(states_2002)` states observed at 7 or more locations. 


```{r}
states_2010 =
  brfss_df |>
  filter(year == "2010") |> 
  group_by(locationabbr) |>
  summarize(locations = n_distinct(locationdesc)) |>
  filter(locations >= 7) 
kable(states_2010, col.names = c("States", "Number of Locations"))
```
In 2010, there are `r nrow(states_2010)` states observed at 7 or more locations.


```{r}
excl_brfss =
  brfss_df |>
  filter(response == "Excellent") |>
  group_by(year, locationabbr) |>
  mutate(avg_data = mean(data_value)) |>
  select(year, locationabbr, avg_data)

#Create the plot
ggplot(excl_brfss, aes(x = year,
                       y = avg_data,
                       group = locationabbr,
                       color = locationabbr)) +
  geom_line(size = 1) + 
  labs(title = "Average Data Value Over Time within a State (Excellent)",
       x = "Year",
       y = "Average Data Value",
       color = "State") +
  theme_minimal()
```
This spaghetti plot shows the average data value over time within each state, limited to the response "Excellent". There are a total of `r length(unique(excl_brfss$locationabbr))` states observed. To make the plot more readable, the legends indicating `State` have been placed on the side of the plot, with the x-axis representing the `Year`, and the y-axis representing the `Mean Value`.


Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State
```{r}
brfss_df |>
  filter(locationabbr == "NY", year %in% c(2006, 2010)) |>
  separate(locationdesc, into = c("state", "county"), "-") |>
  select(year, county, response, data_value) |>
  ggplot(aes(x = response, y = data_value)) +
  geom_boxplot() +
  geom_point(aes(color = county)) +
  facet_grid(~ year) +
  labs(title = "Distribution of Data Value for Responses in New York State (2006 & 2010)",
       x = "Response",
       y = "Data Value",
       color = "County") +
  theme_minimal()
```
comment:

## Problem 3
Your final dataset should include all originally observed variables; exclude participants less than 21 years of age, and those with missing demographic data; and encode data with reasonable variable classes (i.e. not numeric, and using factors with the ordering of tables and plots in mind).
```{r}
#Import and tidy covariate info
covar_df =
  read.csv("data/nhanes_covar.csv", skip = 4) |>
  janitor::clean_names() |>
  mutate(
    sex = recode(sex, "1" = "male", "2" = "female"),
    education = recode(education, "1" = "Less than high school", "2" = "High school equivalent", "3" = "More than high school")) |>
  mutate(
    sex = factor(sex), education = factor(education)) |>
  filter(age >= 21) |>
  drop_na()

#Import and tidy accelerometer info
accel_df =
  read.csv("data/nhanes_accel.csv") |>
  janitor::clean_names()

#combine two datasets
merged_nhanes = merge(covar_df, accel_df, by = "seqn")
```


Table for the number of men and women in each education category
```{r}
gender_edu = merged_nhanes |>
  group_by(sex, education) |>
  summarize(count = n(), .groups = "drop") |>
  pivot_wider(names_from = sex, values_from = count)
kable(gender_edu,
      title = "Number of Men and Women in each Education Category")
```

Visualization of the age distributions for men and women in each education category.
```{r}
ggplot(merged_nhanes, aes(x = education, y = age, fill = sex)) +
  geom_boxplot(alpha = 0.3) +
  labs(title = "Age Distributions by Gender and Education",
       x = "Education level",
       y = "Age") +
  theme_minimal()
```


Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate across minutes to create a total activity variable for each participant. Plot these total activities (y-axis) against age (x-axis); your plot should compare men to women and have separate panels for each education level. Include a trend line or a smooth to illustrate differences. Comment on your plot.
```{r}
plot2_nhanes = merged_nhanes |>
  mutate(total_activity = rowSums(select(merged_nhanes, starts_with("min"))))

#Make a plot
ggplot(plot2_nhanes, aes(x = age, y = total_activity, color = sex)) +
  geom_point(alpha = 0.5) +
  facet_wrap(~education) +
  geom_smooth() +
  labs(title = "Total Activity by Age, Gender, Education",
       y = "Total Activity",
       x = "Age")
```


Accelerometer data allows the inspection activity over the course of the day. Make a three-panel plot that shows the 24-hour activity time courses for each education level and use color to indicate sex. Describe in words any patterns or conclusions you can make based on this graph; including smooth trends may help identify differences.

```{r}
plot3_nhanes = merged_nhanes |>
  group_by(education, sex) |>
  summarize(across(starts_with("min"), ~mean(.), .names = "mean_{.col}")) |>
  pivot_longer(cols = starts_with("mean_"), names_to = "time", values_to = "mean") |>
  mutate(time = substring(time, 9),
         time = as.numeric(time))
  
  
ggplot(plot3_nhanes, aes(x = time, y = mean, color = sex)) +
  geom_point(alpha = 0.1) +
  facet_wrap(~education) +
  geom_smooth() +
  labs(title = "24-hour Activity by Age and Education",
       y = "Mean Value of Activity",
       x = "Time(min)",
       color = "Sex")
```

